Static Race Detection in OS Kernels by Mining Locking Rules

### 2023-07-27 3.technique.tex
Finish the draft of technique

\subsection{Pattern-Based Estimation}
\label{subsec_estimation}

>>>>

\subsection{Pattern-Based Estimation}
\label{subsec_estimation}
Many data races are benign or introduced by developers deliberately to improve 
the kernel performance. They can not cause memory or logic bugs, and thus 
developers are unwilling to put effort into repairing them. We observe that 
harmful data races match some typical patterns, and propose a {\em 
pattern-based estimation} to extract data races that can cause memory or logic 
bugs.

At present, we propose four patterns that can cause null-pointer dereference, 
infinite loop, data inconsistency and unprotected write, because these four 
patterns are common and dangerous in OSes. The four patterns are shown in 
Figure~\ref{fig_pattern}, and we assume that accesses to the fields of {\em 
dev} should be protected by {\em dev->lock}.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.85\linewidth]{figures/fig_pattern.pdf}
	\figcaption{Three harmful patterns of data races.}
	\label{fig_pattern}
\end{figure}

\begin{itemize}
	\item \PP{Null-pointer dereference.} The two accesses to {\em dev->event} 
	are not protected by {\em dev->lock}. This can cause a null-pointer 
	dereference if {\em dev->event} is set to NULL right after the condition of 
	the if statement is checked to be true. To recognize such pattern, the 
	analysis first locates the data race, if it is checked by an if statement 
	and then dereferenced, a possible null-pointer dereference can occur.
	\item \PP{Infinite loop.} The condition of the while statement does not 
	protected by {\em dev->lock}. This can cause a infinite loop when {\em 
	dev->cnt} is assigned with a non-zero value every time before it is checked 
	by the while statement. The analysis detects such bugs by judge whether the 
	data race acts as a condition of a loop statement.
	\item \PP{Data inconsistency.} Two different fields of the same data 
	structure are accessed without acquiring {\em dev->lock}. This can cause a 
	data inconsistency when the data structure field {\em dev->clock} is 
	changed by another thread right after the access to {\em dev->clock->min}. 
	To recognize such pattern, the analysis detect whether multiple fields of 
	the same data structure are accessed without acquiring the protecting lock.
	\item \PP{Unprotected write.} The write access to {\em dev->event} is not 
	protected by {\em dev->lock}. This is dangerous because the value {\em 
	dev->event} can be modified at any time when other threads access it, and 
	thus introduce unpredictable behavior. The analysis detect such pattern by 
	judge whether a data race occurs in a write access.
\end{itemize}
